---
title: "Petrogale_climate"
author: "D. Rex Mitchell"
date: "2024-06-09"
output: html_document
editor_options: 
  chunk_output_type: console
---

Load packages
```{r}
library(geomorph)
library(landvR)
library(vegan)
library(phytools)
library(usdm)
```

Import data
```{r}
load("Petrogale_climate_tests.rda")
```

Perform Procrustes superimpositions and extract centroid size
```{r}
ind <- dimnames(shapedata)[[3]]
pet_sym <- bilat.symmetry(shapedata, ind=ind, land.pairs = sym, object.sym=T, curves=slide)
gpa <- gpagen(shapedata)
size <- gpa$Csize
```

An initial simple test of Bergmann's rule, regression cranial size and latitude
```{r}
bergmann.model <- procD.lm(log(size)~lat, iter=999); summary(bergmann.model)
```

Prepare data of VARPART analysis
```{r}
geodata <- cbind(lon, lat)
distances <- dist(geodata) # generates pairwise distances matrix
spatial.data<-pcnm(distances) # Principal Coordinate Neighbour Matrices conversion

length(spatial.data$values)
sum(spatial.data$values[1:8]/sum(spatial.data$values[1:152])) # Identify how many values equate to ~95% and isolate these
spat <- spatial.data$vectors[,1:8]

vif <- vifstep(climdata,th=3); vif 
# 5 variables remain after removing correlated climate variables
```

Do VARPART analysis
```{r}
sym.mat <- two.d.array(pet_sym$symm.shape)
mod<-varpart(sym.mat,~log(size),~spat,~climdata[,c(2:3,8:9,18)]); mod
showvarparts(3)
plot(mod,digits = 2, Xnames = c('cranial size','geography', 'climate'), bg = c('red','orange', 'blue'))
```

Check R-squared values not printed with the model
```{r}
full_model <- rda(sym.mat ~ log(size) + spat + climdata[,c(2:3,8:9,18)])
anova(full_model)
RsquareAdj(full_model)

full_size_fraction <- rda(sym.mat ~ log(gpa$Csize))
anova(full_size_fraction) 
RsquareAdj(full_size_fraction)

full_geo_fraction <- rda(sym.mat ~ spat)
anova(full_geo_fraction) 
RsquareAdj(full_geo_fraction)

full_clim_fraction <- rda(sym.mat ~ climdata[,c(2:3,8:9,18)])
anova(full_clim_fraction) 
RsquareAdj(full_clim_fraction)

pure_size_fraction <- rda(sym.mat ~ log(size) + Condition(climdata[,c(2:3,8:9,18)]) + Condition(spat))
anova(pure_size_fraction)
RsquareAdj(pure_size_fraction)

pure_geo_fraction <- rda(sym.mat~ spat + Condition(log(size)) + Condition(climdata[,c(2:3,8:9,18)]))
anova(pure_geo_fraction)
RsquareAdj(pure_geo_fraction)

pure_clim_fraction <- rda(sym.mat ~ climdata[,c(2:3,8:9,18)] + Condition(log(size)) + Condition(spat))
anova(pure_clim_fraction)
RsquareAdj(pure_clim_fraction)
```

Latitude and climate models, with predicted shapes
```{r}
latitude_shape <- procD.lm(pet_sym$symm.shape~lat,iter=999); summary(latitude_shape)
latitude_plot <- plot(latitude_shape, type = "regression", predictor = lat, reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species) 

plot(latitude_plot$RegScore~lat, pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Latitude", ylab = "Shape Score")

preds <- shape.predictor(latitude_shape$GM$fitted, x = latitude_plot$PredLine,
                         predmin = max(latitude_plot$PredLine), 
                         predmax = min(latitude_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)

MDR <- procD.lm(pet_sym$symm.shape~climdata[,2], iter=999); summary(MDR)
MDR_plot <- plot(MDR, type = "regression", predictor = climdata[,2], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species) 

plot(MDR_plot$RegScore~climdata[,2], pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Mean Diurnal Range", ylab = "Shape Score")

MDR_size <- procD.lm(MDR_plot$RegScore~log(size), iter=999); summary(MDR_size)

preds <- shape.predictor(MDR$GM$fitted, x = MDR_plot$PredLine, 
                         predmin = min(MDR_plot$PredLine), 
                         predmax = max(MDR_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)

ISO <- procD.lm(pet_sym$symm.shape~climdata[,3], iter=999); summary(ISO)
ISO_plot <- plot(ISO, type = "regression", predictor = climdata[,3], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species) 

plot(ISO_plot$RegScore~climdata[,3], pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Isothermality", ylab = "Shape Score")

ISO_size <- procD.lm(ISO_plot$RegScore~log(size), iter=999); summary(ISO_size)
preds <- shape.predictor(ISO$GM$fitted, x = ISO_plot$PredLine, 
                         predmin = min(ISO_plot$PredLine), 
                         predmax = max(ISO_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)

TWQ <- procD.lm(pet_sym$symm.shape~climdata[,8], iter=999); summary(TWQ)
TWQ_plot <- plot(TWQ, type = "regression", predictor = climdata[,8], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species) 

plot(TWQ_plot$RegScore~climdata[,8], pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Temperature of Wettest Quarter", ylab = "Shape Score")

TWQ_size <- procD.lm(TWQ_plot$RegScore~log(size), iter=999); summary(TWQ_size)
preds <- shape.predictor(TWQ$GM$fitted, x = TWQ_plot$PredLine, 
                         predmin = min(TWQ_plot$PredLine), 
                         predmax = max(TWQ_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)

TDQ <- procD.lm(pet_sym$symm.shape~climdata[,9], iter=999); summary(TDQ)
TDQ_plot <- plot(TDQ, type = "regression", predictor = climdata[,9], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species)

plot(TDQ_plot$RegScore~climdata[,9], pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Temperature of Driest Quarter", ylab = "Shape Score")

TDQ_size <- procD.lm(TDQ_plot$RegScore~log(size), iter=999); summary(TDQ_size)
preds <- shape.predictor(TDQ$GM$fitted, x = TDQ_plot$PredLine, 
                         predmin = min(TDQ_plot$PredLine), 
                         predmax = max(TDQ_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)

PWQ <- procD.lm(pet_sym$symm.shape~climdata[,18], iter=999); summary(PWQ)
PWQ_plot <- plot(PWQ, type = "regression", predictor = climdata[,18], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species)

plot(PWQ_plot$RegScore~climdata[,18], pch=icons, cex=exp(gpa$Csize/100)/40, bg = species, xlab = "Precipitation of Warmest Quarter", ylab = "Shape Score")

PWQ_size <- procD.lm(PWQ_plot$RegScore~log(size), iter=999); summary(PWQ_size)
preds <- shape.predictor(PWQ$GM$fitted, x = PWQ_plot$PredLine, 
                         predmin = min(PWQ_plot$PredLine), 
                         predmax = max(PWQ_plot$PredLine))
differences <- coordinates.difference(coordinates = preds$predmax,
                                      reference = preds$predmin, type = "spherical")
procrustes.var.plot(preds$predmin, preds$predmax, col = heat.colors, col.val = differences[[1]][1:150], magnitude = 1, pt.size = 1)
```

Individual tests of shape relative to separate climate variables
```{r}

lms <- list()
predplot <- list()
lms_size <- list()

par(mfrow=c(10,2))

for (i in 1:ncol(climdata)){
  #make lm of shape vs climate data
  temp_lm <- procD.lm( pet_sym$symm.shape ~ climdata[,i], iter=999)
  #store the summary for tabulating later
  lms[[i]]<- summary(temp_lm)
  #Create plots for plotting and stats of regcores below
  predplot[[i]] <- plot(temp_lm, type = "regression", predictor = climdata[,i], reg.type = "RegScore", pch=icons, cex=gpa$Csize/200, bg = species)
  #regscore vs csize
  lms_size[[i]] <- summary(procD.lm(predplot[[i]]$RegScore~log(size), iter=999))
  
}

#make table that has as many rows as variables and columns as numbers of summary stats columns
shape_summary_table <- matrix (nrow = ncol(climdata), ncol = ncol(lms[[1]]$table))
rownames(shape_summary_table) <- colnames(climdata)
colnames(shape_summary_table) <- colnames(lms[[1]]$table)

#1st row is association of shape with climate
for (i in 1:ncol(climdata)) {
  
  shape_summary_table[i,] <- as.matrix(lms[[i]]$table)[1,]
  
}

shape_summary_table [,2:6] <- round (shape_summary_table [,2:6], 2)

#make table for SIZE that has as many rows as variables and columns as numbers of summary stats columns
size_summary_table <- matrix (nrow = ncol(climdata), ncol = ncol(lms_size[[1]]$table))
rownames(size_summary_table) <- colnames(climdata)
colnames(size_summary_table) <- colnames(lms_size[[1]]$table)

#1st row is association of shape with climate
for (i in 1:ncol(climdata)) {
  
  size_summary_table[i,] <- as.matrix(lms_size[[i]]$table)[1,]
  
}

size_summary_table [,2:6] <- round (size_summary_table [,2:6], 2)

Full_table <- cbind (shape_summary_table, size_summary_table)

write.csv (Full_table, file="Individual_climate_variables.csv")







```

Phylogenetic tests compared with OLS MANOVA models
```{r}
OLS_allom <- procD.lm(pet_sym$symm.shape ~ log(size), iter=999); summary(OLS_allom)

PGLS_allom <- procD.pgls(meanshapes_ord ~ log(size_ord), tree, iter=999); summary(PGLS_allom)

OLS_clim <- procD.lm(pet_sym$symm.shape ~ climdata[,18], iter=999); summary(OLS_clim)

PGLS_clim <- procD.pgls(meanshapes_ord ~ clim_ord, tree, iter=999); summary(PGLS_clim)

PGLS_sizeclim <- procD.pgls(log(size_ord) ~ clim_ord, tree, iter=999); summary(PGLS_sizeclim)

# Test for phylogenetic signal of size and BIO18
physignal(log(size_ord), tree, iter=999)
physignal(clim_ord, tree, iter=999)
```

Phylogenetic trees coloured by size and BIO18.
```{r}
plot(tree)
axisPhylo()

phyplot_size <- contMap(tree, size_ord)
phyplot_clim <- contMap(tree, clim_ord)
```